A Large-Scale Study of the Use of Eval in R

### 2020-08-28 inst/paper/main.tex
patterns

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{hyperref}
\newcommand{\missingTag}[1]{\textcolor{red}{#1}\xspace}
\newcommand{\missingNumber}{\textcolor{red}{XX}\xspace}
\newcommand{\missingPercentage}{\textcolor{red}{XX\%}\xspace}
\newcommand{\missingTable}{\textcolor{red}{XXTable}\xspace}
\newcommand{\missingGraph}{\textcolor{red}{XXGraph}\xspace}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage{xspace}
\usepackage{url}

@@

%%In package \emph{gamlss::get.K} function \c{get.K} evaluates its expression in its
%%parent environment, the one of the body of \c{get.K} returned from
%%\c{environment()}.
%%
%%\begin{lstlisting}
%%eval(parse(text = paste(\"fam$dld\", ii, sep = \"\")), envir = environment())
%%\end{lstlisting}
%%
%%
%%The \emph{glue} library is used to interpolate string litterals with variables, such as in \c{glue('My name is \{name\}.')} where \c{name} will be subtituted by its value. \emph{glue} makes it possible to define transformers that process those interpolated variables. They are built upon the \c{identity_transformer}, which uses \eval. The interpretation environment for \eval is often global here, as the string litterals are often used to print variables at the top level.
%%
%%\begin{lstlisting}
%%identity_transformer <- function(text, envir) 
%%{
%%    eval(parse(text = text, keep.source = FALSE), envir)
%%}
%%\end{lstlisting}
%%
%%

@@

This section presents the common patterns of \eval usage in our corpus. We
randomly selected 20 packages from the corpus and studied how they use eval.
We have grouped those uses in the following categories.

@@

  \subsubsection{Miscellaneous} \eval is also used in R packages to get around
  various limitations. For example, in \datatable package, \eval is used to modify
  a binding in the \base package of R. Bindings in the \base package are locked
  for modification and require explicit unlocking using \unlockBinding function.
  However, directly using this function triggers a warning during package building
  process which statically analyzes the package for use of such restricted
  functions. The \datatable gets around this limitation by enclosing the call to
  \unlockBinding inside \eval.

>>>>

\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{listings}
\usepackage{hyperref}
\newcommand{\missingTag}[1]{\textcolor{red}{#1}\xspace}
\newcommand{\missingNumber}{\textcolor{red}{XX}\xspace}
\newcommand{\missingPercentage}{\textcolor{red}{XX\%}\xspace}
\newcommand{\missingTable}[1][XX]{\textcolor{red}{Table #1}\xspace}
\newcommand{\missingGraph}{\textcolor{red}{XXGraph}\xspace}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{paralist}
\usepackage{xspace}
\usepackage{url}

@@

%%In package \emph{gamlss::get.K} function \c{get.K} evaluates its expression in its
%%parent environment, the one of the body of \c{get.K} returned from
%%\c{environment()}.
%%
%%\begin{lstlisting}
%%eval(parse(text = paste(\"fam$dld\", ii, sep = \"\")), envir = environment())
%%\end{lstlisting}
%%
%%
%%The \emph{glue} library is used to interpolate string litterals with variables, such as in \c{glue('My name is \{name\}.')} where \c{name} will be subtituted by its value. \emph{glue} makes it possible to define transformers that process those interpolated variables. They are built upon the \c{identity_transformer}, which uses \eval. The interpretation environment for \eval is often global here, as the string litterals are often used to print variables at the top level.
%%
%%\begin{lstlisting}
%%identity_transformer <- function(text, envir) 
%%{
%%    eval(parse(text = text, keep.source = FALSE), envir)
%%}
%%\end{lstlisting}
%%
%%
%% \end{lstlisting}
%%$ <- to fix emacs fontification issues

@@

The JavaScript paper identified a number of frequently occurring eval
patterns. The patterns were detected using simple regular expressions on the
string passed to \eval. In the case of R look at the AST passed to evals.

\missingTable[JS evals in R]

For the rest of this section we show some of the common use of \eval in R that
we have seen by looking at 20 randomly selected packages:

@@

  \subsubsection{Miscellaneous} \eval is also used in R packages to get around
  various limitations. For example, in \datatable package, \eval is used to
  modify a binding in the \base package of R. Bindings in the \base package are
  locked for modification and require explicit unlocking using \unlockBinding
  function. However, directly using this function triggers a
  restricted-function-use warning during package build, preventing the package
  to be accepted by CRAN. The \datatable gets around this limitation by
  enclosing the call to \unlockBinding inside \eval. From the same reason, other
  packages use \eval to access private variables and functions from their
  dependencies.
  % \AllValueInputEvalCallPerc migth hit on a wong usage
